Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-08 paper/v2/paper-edited.tex
more edits

\sysname allows operators to configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the 
different routes traffic may (or may not) take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that for routes leading to our customer,
we prefer using \CD{R1} over \CD{R2} (policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.  
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that 
leave through \CD{Peer} and so then \CD{Prov}.  As we
describe in the next section, the \Exit\ predicate, as well as other
path predicates used later in this section, is simply a shorthand for 
a particular regular expression over paths that is expressible in our policy
language.

% DAVE CONVERSIONS
% C_{pfx} ===> PCust
% I_{pfx} ===> PInternal

To associate ranked paths with
one or more prefixes, we define a \sysname \emph{policy}.
Within a policy, statements with the form $t$\Path$p$
associate the prefixes defined by the predicate $t$ with the set of
ranked paths defined by the path expression $p$.  In general,
prefix predicates can be defined by arbitrary boolean combinations
(and, or, not) of concrete prefixes and community tags.  Here,
we assume we have already defined predicates \CD{PCust} for the
customer prefixes.
%and \CD{PInt} for our internal prefixes.
In the following code, ranked paths are associated with
customer prefixes,
and all other prefixes (\CD{true}).\footnote{Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements.  Hence, when the prefix predicate \True{} follows
the statement involving \CD{PCust}, it is interpreted as
\CD{\True{} \AND{} \NOT{}PCust}.}

@@

Line 2 of this policy 
restricts traffic destined to known customer prefixes (\CD{PCust}) to only follow paths that end at the customer. In addition, it enforces the network-wide preference that traffic leaves through \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}.
%defines the paths that exit our network to customer prefixes---\CD{ExitCust} defines paths through \CD{R1} and \CD{R2}, and in the event
%that connections to the customer through both routers fail,
%a backup route (\CD{\End(Cust)}) is defined that admits traffic along
%any path (including those through other networks) that ends at the customer.
%Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.
Line 3 applies to any other traffic, and allows the traffic to leave through any direct neighbor with the usual preferences of \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}. To summarize our progress, the \CD{Routing} policy
implements P1, P3, and P4 from Figure~\ref{fig:example1}.

Since by default routing allows transit traffic (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints. 
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X \OR Y) \AND \Exit(X \OR Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between $X$ and $Y$ as traffic that enters the network through $X$ or $Y$ and then leaves the network through $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

\begin{code}
\Define NoTrans = 
  \{\True \Path \NOT{}transit(Peer,Prov) \AND 
           \NOT{}cust-transit(Cust,Prov)\}
\end{code}
\noindent
The \CD{NoTrans} constraint above requires that all traffic not follow a path that transits the network between \CD{Peer} and \CD{Prov}. Additionally, it prevents traffic from ever following paths that leave our network and later pass through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
and \CD{NoTrans} simultaneously, we simply conjoin them: \CD{Routing \AND{} NoTrans}.
%
%% \begin{lstlisting}[mathescape=true]
%% Routing $\AND$  NoTransit
%% \end{lstlisting}
%
These constraints capture the entire policy. From them, \sysname will generate per-device import and export filters, local preferences,
MED attributes, and community tags to ensure that the policy is
implemented correctly under all failure cases.

@@

\begin{code}
\Define Ownership = 
    \{PG1 \Path \End(A)
     PG2 \Path \End(B)
     PL1 \Path \End(E)
     PL2 \Path \End(F)
     \True \Path \End(\Out)\}
\end{code}
\noindent
In English: traffic for prefix \CD{PG1} is only allowed to follow paths that
end at router \CD{A}; traffic matching \CD{PG2},
but not \CD{PG1}, must
end at router \CD{B}; and so on.
Any traffic destined for a prefix that not a part of the data center should be allowed to leave the data center and end at some external location.
The special keyword \Out{} represents any location outside the user's network, and the keyword \In{} represents any location inside the network.
For the second constraint, we can define another policy:

@@

As in the backbone example, we can logically conjoin these constraints
to specify the network-wide policy.
In addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane.
For instance, a constraint on aggregation is included to ensure that
aggregation for global prefixes is performed from
locations inside (\In) the network to locations outside (\Out). 
In this case, \CD{PG1} and \CD{PG2} will use the aggregate \CD{PG} 
(which we assume is defined earlier using an appropriate prefix such 
as \CD{74.3.28.0/24})
when advertised outside the datacenter.

@@

As before, once \sysname compiles the policy, it is guaranteed to be policy-compliant under all possible failure scenarios. In addition, \sysname will efficiently find a lower bound on the number of failures required to create an aggregation-induced black hole.
%It can efficiently determine a lower bound on the number of failures necessary to black hole traffic given the user's policy, topology, and placement of specific aggregates.

@@

We handle these challenges by breaking compilation into multiple stages and developing efficient algorithms for each (See Figure~\ref{fig:pipeline}). The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines that the PGIR is safe, it translates it to an abstract BGP (ABGP) representation. Vendor-specific adaptors to translate from ABGP to concrete device configurations or to integrate with an existing template-based system are easily added. The \sysname compiler currently generates Quagga configurations by default.

@@

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& a.b.c.d/n & \textit{prefix} \\
     c       &::=& \cdots & \textit{community value} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = c & \textit{community test} \\
     r       &::=& n & \textit{AS number} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT(r) & \textit{path negation} \\
         &\BNFALT& r^* & \textit{repetition} \\
     l       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, l) \BNFALT tag(c, t, l) & \textit{control constraints} \\
  \end{array} \]

@@

\para{Syntax}
The \sysname FE is just a thin layer atop a core, regular-expression language (RIR) for describing preference-based path constraints. 
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which have a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use other standard abbreviations
for regular expressions such as $r^+$, which is simply syntactic sugar for $r \cdot r^*$, a sequence of one or more occurrences of $r$, when appropriate.
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size and churn in practice. 
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.

>>>>

\sysname allows operators to configure the network with the
abstraction that they have centralized control over routing.
Specifically, the operator simply provides a set of high-level constraints
that describe the 
different routes traffic may (or may not) take and their relative preferences.
\sysname specifications are written modularly via a series
of declarations.
%These definitions allow users to express the three
%major elements of any \sysname specification:  \emph{prefixes},
%\emph{paths} and \emph{policies}.
For example, to begin specification of the backbone
network from the previous section, we first express the idea that 
we prefer that traffic leave the network through \CD{R1} over \CD{R2} (to \CD{Cust}) over \CD{Peer} over \CD{Prov} (policy P3 from Figure~\ref{fig:example1}):
%\begin{lstlisting}[mathescape=true]
%define ExitCust = $exit(R_1 \gg R_2)$
%\end{lstlisting}
\begin{code}
\Define Prefs = \Exit(R1 \Prefer R2 \Prefer Peer \Prefer Prov)
\end{code}
This  statement defines a set of \emph{ranked paths}, which includes
all paths (and only those paths) for which traffic exits our network
through either router \CD{R1}, router \CD{R2}, \CD{Peer}, or \CD{Prov}.  
The paths that exit through \CD{R1}
are preferred (\Prefer) to those that exit through \CD{R2}, which are preferred to those that 
leave through \CD{Peer} and then \CD{Prov}.  As we
describe in the next section, the \Exit\ predicate, as well as other
path predicates used later in this section, is simply a shorthand for 
a particular regular expression over paths that is expressible in our policy
language. The preference operator (\Prefer) is flexible and can be used between constraints as well as among individual routers. For example, the above constraint could have been written equivalently as 
\Exit(\CD{R1}) \Prefer \ldots \Prefer ~\Exit(\CD{Prov})

To associate ranked paths with
one or more prefixes, we define a \sysname \emph{policy}.
Within a policy, statements with the form $t$\Path$p$
associate the prefixes defined by the predicate $t$ with the set of
ranked paths defined by the path expression $p$.  In general,
prefix predicates can be defined by arbitrary boolean combinations
(and, or, not) of concrete prefixes and community tags.  Here,
we assume we have already defined predicates \CD{PCust} for the
customer prefixes.
%and \CD{PInt} for our internal prefixes.
In the following code, ranked paths are associated with
customer prefixes,
and all other prefixes (\CD{true}). Policy statements are processed in
order with earlier policy statements taking precedence over later
policy statements. Hence, when the predicate \True{} follows
the statement involving \CD{PCust}, it is interpreted as
\CD{\True{} \AND{} \NOT{}PCust}.

@@

Line 2 of this policy 
restricts traffic destined to known customer prefixes (\CD{PCust}) to only follow paths that end at the customer. In addition, it enforces the network-wide preference that traffic leaves through \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}.
%defines the paths that exit our network to customer prefixes---\CD{ExitCust} defines paths through \CD{R1} and \CD{R2}, and in the event
%that connections to the customer through both routers fail,
%a backup route (\CD{\End(Cust)}) is defined that admits traffic along
%any path (including those through other networks) that ends at the customer.
%Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.
Line 3 applies to any other traffic not matching \CD{PCust} and allows the traffic to leave through any direct neighbor with the usual preferences of \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}. To summarize our progress, the \CD{Routing} policy
implements P1, P3, and P4 from Figure~\ref{fig:example1}.

Since by default routing allows transit traffic (\EG, traffic entering from
\CD{Peer} and leaving through \CD{Prov}), we separately define a
policy to enforce P2 and P5 from Figure~\ref{fig:example1}, using conjunction (\AND),
disjunction (\OR) and negation (\NOT) of constraints. 
First, we create reusable abstractions for describing traffic that transits our network. In \sysname, this is done by creating a new parameterized definition.
\begin{code}
\Define transit(X,Y) = \Enter(X \OR Y) \AND \Exit(X \OR Y)
\Define cust-transit(X,Y) = \Later(X) \AND \Later(Y)
\end{code}
\noindent
Here we define transit traffic between groups of neighbors $X$ and $Y$ as traffic that enters the network through some neighbor in $X$ or $Y$ and then leaves the network through some neighbor in either $X$ or $Y$. Similarly, we define customer transit for customer $X$ and provider $Y$ as traffic that later goes through both $X$ and $Y$ after leaving our network. Using these two new abstractions, we can now implement policies P2 and P5 with the following constraint.

\begin{code}
\Define NoTrans = 
  \{\True \Path \NOT{}transit(Peer,Prov) \AND 
           \NOT{}cust-transit(Cust,Prov)\}
\end{code}
\noindent
The \CD{NoTrans} constraint above requires that all traffic not follow a path that transits the network between \CD{Peer} and \CD{Prov}.
Additionally, it prevents traffic from ever following paths that leave our network and later go through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
and \CD{NoTrans} simultaneously, we simply conjoin them: \CD{Routing \AND{} NoTrans}.
%
%% \begin{lstlisting}[mathescape=true]
%% Routing $\AND$  NoTransit
%% \end{lstlisting}
%
These constraints capture the entire policy. From them, \sysname will generate per-device import and export filters, local preferences,
MED attributes, and community tags to ensure that the policy is
implemented correctly under all failures.

@@

\begin{code}
\Define Ownership = 
    \{PG1 \Path \End(A)
     PG2 \Path \End(B)
     PL1 \Path \End(E)
     PL2 \Path \End(F)
     \True \Path \End(\Out)\}
\end{code}
\noindent
In English: traffic for prefix \CD{PG1} is only allowed to follow paths that
end at router \CD{A}; traffic matching \CD{PG2},
but not \CD{PG1}, must
end at router \CD{B}; and so on.
Any traffic destined for a prefix that not a part of the data center should be allowed to leave the data center and end at some external location, which is otherwise unconstrained.
The special keyword \Out{} here matches any location outside the user's network, while the keyword \In{} will match any location inside the network.
For the second constraint, we can define another policy:

@@

As in the backbone example, we can logically conjoin these constraints
to specify the network-wide policy.
However, in addition to constraints on the shape of paths,
\sysname allows the operator to specify constraints on the BGP control plane itself.
For instance, a constraint on aggregation is included to ensure that
aggregation for global prefixes is performed from
locations inside (\In) the network to locations outside (\Out). 
In this case, \CD{PG1} and \CD{PG2} will use the aggregate \CD{PG} 
(which we assume is defined earlier using an appropriate prefix such 
as \CD{74.3.28.0/24})
when advertised outside the datacenter.

@@

Once \sysname compiles the policy, it is guaranteed to be compliant under all possible failure scenarios, modulo aggregation black holes (which are unavoidable in general). In the presence of aggregation, the \sysname compiler will also efficiently find a lower bound on the number of failures required to create an aggregation-induced black hole.

@@

We handle these challenges by breaking compilation into multiple stages and developing efficient algorithms for each (See Figure~\ref{fig:pipeline}). The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines that the PGIR is safe, it translates it to an abstract BGP (ABGP) representation. Vendor-specific adaptors to translate from ABGP to concrete device configurations or to integrate with an existing template-based system are easily added. The \sysname compiler currently generates Quagga router configurations by default.

@@

     pol     &::=& p_1, \dots, p_n & \textit{constraints} \\
     p       &::=& t \hspace{.3em} \Path \hspace{.3em} r_1 \Prefer \dots \Prefer r_m & \textit{preferences} \\
     x       &::=& d.d.d.d/d & \textit{prefix} \\
     t       &::=& \True & \textit{true} \\
         &\BNFALT& \NOT t & \textit{negation} \\
         &\BNFALT& t_1 \OR{} t_2 & \textit{disjunction} \\
         &\BNFALT& t_1 \AND{} t_2 & \textit{conjunction} \\
         &\BNFALT& \textit{prefix} = x & \textit{prefix test} \\
         &\BNFALT& \textit{comm} = d & \textit{community test} \\
     r       &::=& n & \textit{AS number} \\
         &\BNFALT& \emptyset & \textit{empty set} \\
         &\BNFALT& \In & \textit{internal loc} \\
         &\BNFALT& \Out & \textit{external loc} \\
         &\BNFALT& r_1 \cup r_2 & \textit{union} \\
         &\BNFALT& r_1 \cap r_2 & \textit{intersection} \\
         &\BNFALT& r_1 \cdot r_2 & \textit{concatenation} \\
         &\BNFALT& \NOT r & \textit{path negation} \\
         &\BNFALT& r^* & \textit{iteration} \\
     l       &::=& r_1 \rightarrow r_2 & \textit{link pairs} \\
     cc     &::=& agg(x, l) \BNFALT tag(c, t, l) & \textit{control constraints} \\
  \end{array} \]

@@

\para{Syntax}
The \sysname FE is just a thin layer atop a core, regular-expression language (RIR) for describing preference-based path constraints. 
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which have a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expressions abbreviation $r^+$ to stand for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path). 
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

\sysname also supports constraints purely on the control-plane behavior of BGP. For example, prefix aggregation is an important optimization to reduce routing table size and churn in practice. A constraint of the form $agg(x,l)$ tells the compiler to perform aggregation for prefix $x$ across all links described by $l$. It is also often useful to be able to add community tags to exported routes in BGP (e.g., to communicate non-standard information to peers). A constraint of the form $tag(c,t,l)$ adds community tag $c$ for any prefixes matching $t$ across links $l$.
%% Aggregation, for example, from internal to external locations, is specified using the same regular syntax as before:
%% $$\Agg(128.17.0.0/16, \In \Link \Out)$$
%% where the expression $\In \Link Out$ refers to control messages flowing from any internal to any external location.
We list only the route aggregation and community tagging constraints in Figure~\ref{fig:rir-syntax}, but we also support other constraints such as limiting the maximum number of routes allowed between ASes, or enabling BGP multipath.
